\bf Visualizing Dynamic Social Network Models

### 2018-10-15 paper-child.Rnw
ready to start submitting....

%Every good \emph{data} analysis includes both numerical and visual summaries of the data, so why restrict \emph{model} analysis, description, and diagnostics to numerical summaries? 
\emph{Model visualization} complements traditional model diagnostic tools such as $R^2$ values and residual plots by enriching our understanding and interpretation of statistical models. Each of the three different meanings of the word ``model", the model \emph{family}, the model \emph{form}, and the \emph{fitted} model \citep[see][pg. 2]{modelvis}, can benefit from model visualization (model-vis). We apply the three principles of model-vis: view the model in the data space, visualize collections of models, and explore the process of fitting the model, to the models and data introduced in Section~\ref{sec:exdatamods}. Since we have already decided on our model family (SAOMs), we shift our focus to the fitted model and the model form. Specifically, we use model visualization to learn more about how the \emph{model form} affects the \emph{fitted model}.

First, we put models in the data space to assess the model fit. %In \citeauthor{modelvis}, they chose define the \textit{data space} as ``the region over which we can reliably make inferences, usually a hypercube containing the data" \citep[p.~206]{modelvis}. But what does this definition mean for dynamic network models?
We define the data space for SAOMs as the combined data structures of the network: the node, edge, and time information. We use the \texttt{R} package \texttt{geomnet} \citep{geomnet} to visualize the three data spaces together. The package lays out the network using node-link diagrams, and the network's node or edge data are then mapped to different visual features.  The color, size, and shape of the points can be used to represent variables in the node data, while the color, linewidth, and linetype of the lines between points can be used to represent the edge variables. To view temporal changes, we place the network at different timepoints side-by-side to see the evolution. By pulling all of this information together using \texttt{geomnet}, we can show the entire data space at once.

<<funsenatevis, fig.height=4, fig.cap='The 111th Senate at two times: while Clinton was in the senate in 2009 (on the left) and after she left (on the right). We map sex, party, and bills authored to the shape, color, and size of the nodes, respectively. We also map the $WPC$, to the edge linewidth.'>>=
se111clint <- filter(seobama, senate == 111)
se111noclint <- filter(seobama, senate == 111)
se111noclint$target[which(se111noclint$target == "Hillary Rodham Clinton")] <- NA
se111clint$Clinton <- 'Yes'
se111noclint$Clinton <- 'No'
clintonSenate <- rbind(se111clint, se111noclint)
clintonSenate$Clinton <- as.factor(clintonSenate$Clinton)
clintonSenate$Clinton <- ordered(clintonSenate$Clinton, levels = c("Yes", "No"))

set.seed(12981120)
clintonSenate %>% filter(!(source %in% c("Roland Burris", "Bernard Sanders", "Frank R. Lautenberg", "Mary L. Landrieu") & is.na(target))) %>%
ggplot() +
  geom_net(directed = T, labelon=F, arrowsize = .3, singletons= F, fiteach = T, arrowgap = .01, layout.alg = 'fruchtermanreingold', ecolour = "grey70",
           aes(from_id = source, to_id = target, color = party, linewidth = gsw, shape = sex,size = n_au/30)) +
  scale_size_continuous(name="# bills\nauthored" , guide = guide_legend(nrow= 2), breaks = c(0,1,2,3), labels = c(0,30, 60, 90)) +
  #theme_net() +
  ThemeNet +
  scale_shape_manual(values = c(17,16), guide = guide_legend(nrow= 2)) +
  scale_color_manual(values = c("black", "grey80","#ef3b2c"), guide = guide_legend(nrow=2)) +
  facet_wrap(~Clinton, nrow = 1, labeller = 'label_both')  +
  theme(legend.position = "bottom")
@

To demonstrate the functionality of \texttt{geomnet}, we visualize the connections in the $111^{th}$ United States Senate at two different timepoints: when Hillary Clinton was in the senate, and after she left to become Secretary of State. Clinton was only in the $111^{th}$ senate from January 3-20, 2009, in the middle of her second term as the junior senator from New York. In that time, she authored two bills and was a cosponsor on 17. There are many senators who cosponsored both of her bills, giving their edges to her a $WPC$ of 1. So with Clinton included in the node-link diagram in Figure~\ref{fig:funsenatevis}, the senate looks much more collaborative than it does without her. We then map our potential covariates to visual features: number of bills authored  maps to the size of the node, gender maps to the shape, and party to the color. In addition, we visualize the strength of the tie by mapping the $WPC$ value between the two senators to the linewidth of the edge. By viewing node information, edge information, and time simultaneously, we have learned that Senator Clinton greatly affected the structure of collaboration within the senate.

<<summnetM1, message=FALSE, fig.cap='On the left, the first wave of observed data that is conditioned on in the model. On the right, the second wave of observed data. In the middle, a summary network from the first model fit to the data. This summary network represents 1,000 simulations of wave 2 using the values from the simple fitted model M1.',results='hide', fig.height=2.5>>=
friendshipData <- array(c(fd2.w1, fd2.w2, fd2.w3), dim = c(16, 16, 3))
friendshipData <- array(c(fd2.w1, fd2.w2, fd2.w3), dim = c(16, 16, 3))
drink1 <- drink[20:35,1]
drink2 <- drink[20:35,2]
drink3 <- drink[20:35,3]
friendship <- sienaDependent(friendshipData)
alcohol <- varCovar(cbind(drink1,drink2, drink3))
mydata <- sienaDataCreate(friendship, alcohol)
M1eff <- getEffects(mydata)
M2eff <- includeEffects(M1eff, jumpXTransTrip, interaction1 = "alcohol")
M3eff <- includeEffects(M1eff, nbrDist2twice)
simu2 <- read.csv("data/simulation-1000-M1-M2-M3.csv")
means <- simu2 %>% group_by(Model, parameter) %>% summarize(
  means = mean(estimate)
)
M1parms <- (means %>% filter(Model == "M1"))$means
M2parms <- (means %>% filter(Model == "M2"))$means
M3parms <- (means %>% filter(Model == "M3"))$means
set.seed(4231352)
#M1sims1000 <- netvizinf::saom_simulate(dat = mydata,
#                            struct = M1eff,
#                       parms = as.numeric(M1parms),
#                       N = 1000)
#save(M1sims1000, file = "data/M1sims1000.RData")

@@

To view a SAOM in the data space, we simulate from the fitted model. No single network instance simulated from a SAOM will fully represent the model, so instead we look at an \textit{expectation} from the network model. We frequently rely on expected values in statistics, but network models, especially those as complex as SAOMs, lack an expected network measure. We could talk about expected values of parameters, but the parameters can be hard to interpret and cannot teach us much about the overall structure of the network. How then, can we arrive at an ``expected" network? We answer this question with visualization.
%Expected values of parameters can be computed, but if they cannot directly tell us anything about our dependent variable, they lose value. 
%There is no defined expected value of a network model, though expectations exist for most other statistical models. 
%For network data, we view an ``average" network through a sort of summary network drawn using the traditional node-link diagram. 
%In Figure~\ref{fig:summnetM1}, we show the \textit{expected network} of wave 2 from model M1, created with 1,000 simulations from the model. 
To construct an expected network, we: 
\begin{enumerate}
\item simulate 1,000 wave 2 and wave 3 instances of our small friendship network from model M1 using parameters that have previously been estimated. Then, we 
\item count the number of times each edge $i \rightarrow j$ appeared in one of the 1,000 network simulations. Then, we
\item construct one weighted summary network for each time point, which has edge weights equal to the proportion of the simulated networks in which the edge appeared. Finally, we 
\item create the \textit{expected network} by including only the edges that appeared in more than 5\% of the simulations. 
\end{enumerate}
The resulting expected network is in the middle panel of Figure~\ref{fig:summnetM1}. Edges with higher weights are darker, thicker lines. On either side of the average network in Figure~\ref{fig:summnetM1}, we show the data for comparison: wave 1 on the left, and wave 2 on the right. We can see that the structure of the average network is much more similar to wave 1 than to wave 2. In both wave 1 and the expected network, the edges form two distinct clusters of nodes: $\{2,3,4,5,6,12,13,15\}$ and $\{7,8,9,10,11,14\}$. The expected network, however, is supposed to represent the second wave of data, on the right in Figure~\ref{fig:summnetM1}. The expected network does not resemble wave 2, which has one small node cluster and on large node cluster, with four unconnected nodes. Because the expected network does not resemble wave 2, we take this as evidence that the simple model, M1, does not capture the change in network structure from the first to the second wave of data. Visualizing the expected network in this way places the model M1 in the data space, and helps us assess model fit. % We can use the average network to help determine model goodness-of-fit. Because the the expected  network looks more like the first wave than the second wave, we can use the visualization in Figure~\ref{fig:summnetM1} as evidence of poor model fit.

@@

Another way to view a SAOM in the data space is with \textit{difference networks} between the expected network and the data, as shown in Figure~\ref{fig:summary2}. We created these networks by computing the difference between the adjacency matrices for the expected network and wave 1, and for wave 2 and the expected network. The lack of model fit becomes even more clear in Figure~\ref{fig:summary2}, with the broken edges representing ties that only show up in the expected network, and the solid edges representing ties in the data that the model misses completely.  Dotted edges are in both the expectation and the data and represent where the model fit the data well. Again, the widths of the edges demonstrate the frequency of the ties in the simulations. On the left in Figure~\ref{fig:summary2}, we see the difference between the first wave of data and the expected network. We only see dotted and broken edges, meaning that the simulated networks consist primarily of edges in wave 1, plus some new edges that are not in the first wave of data. The added ties are all reciprocating edges in the data, meaning that the model predicts more reciprocity between actors than there actually is. On the right in Figure~\ref{fig:summary2}, we see the difference between the second wave of data and the expected network. Because the model is simulating the second wave of data, we would hope to see very few differences between the expected network and wave 2. However, there are six solid edges, which are edges in the data that do not appear at all in the expected network, and there are many broken edges that appear in the expected network but not in the data. Although there are some edges that appear in both the expected network and wave 2, the majority of edges in the expected network are not in wave 2, and there are several edges missing from the expectation. We conclude that the model M1 is a poor fit to our example data, which we have seen very clearly by viewing the model in the data space.

@@

Now that we have assessed model fit, we next want to learn about model behavior in order to choose better fitting models in the future. This is the second principle of model visualization: "collections [of models] are more informative" \citep{modelvis}. We examine the behavior of SAOMs by visualizing collections of models. We build collections of models in four ways. First, we explore the space of all possible models to determine the most significant model parameters.  Second, we vary model settings to uncover relationships between parameters and their effects on the fitted model. Next, we fit the same model to the same data many times to assess the behavior of the underlying simulations. Finally, we fit the same model to different data to learn about fitted parameter values under different conditions. We chose these four collections because they each explore something different about SAOM behavior: the choice of parameters in the objective function, the effects of underlying simulations, and the resulting fitted values under different circumstances. %Using the tools in our visualization toolbox, we view these collections to learn about model behavior.%The first considers the many parameters to include in a SAOM, while the second results in different parameter estimates, so we can see how parameter estimates vary. The third collection shows how those many possible parameters affect the values of parameters in the fitted models. Finally, we explore how the same model behaves under two completely different circumstances.

@@

\noindent \textbf{Exploring the space of all possible models:} The \texttt{RSiena} manual contains over eighty effects to include in the SAOM objective function. To select objective function parameters for fitting models to our example data, we searched through all possible effects available to find \textit{significant} effects. The \texttt{RSiena} package includes the function \texttt{Wald.Rsiena} for Wald-type tests of parameter significance. We start by including the outdegree and reciprocity parameters in the model then add one effect to the objective function, fit the model to the example data, test the added effect for significance, and repeat for all possible model parameters. We performed this procedure, which we used to pick M2 and M3, for both the small friendship and the senate collaboration data. A visualization of the significant effects for the two example data sets are shown in Figure~\ref{fig:sigtesting} in the Appendix. Using this one-at-a-time testing method results in many possible significant parameters of different type and magnitude, demonstrating the flexibility of these models.  %We see in both the friendship data and the senate data results that most of the significant effects have absolute value less than ten. In addition, the $p$-values for the effects from the friendship data are more spread out than the $p$-values for the senate data, which are concentrated at about 0.02 or less. This may suggest that larger data sets tend to result generally in smaller $p$-values, just like a larger sample size results in smaller $p$-values in a $t$-test.

@@

\noindent \textbf{Varying model settings:} The second method for viewing collections of models is varying model settings. We have done this already by choosing three different models M1, M2, M3 to fit to our example data sets. To explore the behavior of these three models, we simulate from them. Using the parameter values given in Table~\ref{tab:meanests} for the friendship data, we simulated 1,000 observations from each of the three models. From the simulations, we created an expected network for wave 2 from models M1, M2, and M3 using the same method described in Section~\ref{sec:minds}. These expected networks are shown in Figure~\ref{fig:averagenet} in the Appendix. We see that all models are failing to capture the structure of the second wave, even though the individual parameters were all significant. %In Section~\ref{sec:samemodsamedat}, we fit these three models to the data 1,000 times, and examine the parameter densities. Here, . For simulation, we used the mean parameter values from the 1,000 fits of the corresponding model.

@@

<<distests, fig.cap="Distribution of fitted parameter values for our three SAOMs. The inclusion of $\\beta_3$ or $\\beta_4$ clearly has an effect on the distribution of the rate parameters, $\\alpha_1$ and $\\alpha_2$, and on the other objective function parameters $\\beta_1, \\beta_2$.", fig.height=3>>=

@@

\noindent \textbf{Fitting the same model to the same data:} The simulations performed to fit a SAOM lead to inherent variability in the fitted model. To examine this variability, we fit models M1, M2, and M3 to the friendship network 1,000 times each. The distributions of the fitted model parameters are shown in Figure~\ref{fig:distests}. Clearly, the inclusion of the JTT parameter, $\beta_3$, results in very different estimates the of the other four parameters in all models, $\alpha_1$, $\alpha_2$, $\beta_1$, and $\beta_2$. On the other hand, the inclusion of the indirect tie parameter, $\beta_4$, does not lead to different estimates of other model parameters. When $\beta_3$ is included, its estimate is postitive, meaning that friendships between two girls with different drinking behaviors tend to form when there is an intermediary who is already friends with the two girls. The inclusion of this parameter leads to increases in the rate parameters' estimates. Thus, when modeling the transitive triplet behavior in this way, we conclude that girls change friends more frequently. It is not clear, however, that the addition of a parameter to the objective function \textit{should} effect the estimates of the rate parameters. The strong relationship between rate parameter estimates and objective function parameters is further shown in Figure~\ref{fig:corplots} in the Appendix. After visualizing the results of the same models fit many times to the same data, we were able to see potential model behavior problems right away.   %In comparison with models M1 and M3, model M2 typically has higher estimates of the rate parameter, meaning that the inclusion of the covariate statistic in the model leads to higher estimates of the number of times, on average, a node gets to change its ties.

@@

% table of fitted M1 M2 M3
\begin{table}
\scalebox{0.9}{
\begin{tabular}{lcccccc }

@@

\noindent \textbf{Fitting the same model to different data:} Finally, we view collections of models by fitting the same models to different data to understand how the model behaves in different contexts. As we did for the friendship data, we estimated the parameters for M1, M2, and M3, when fit to the senate data (see Figure \ref{fig:senateplot}) 100 times each. The means and standard deviations of the repeated parameter estimates for all six model-data combinations are given in Table~\ref{tab:meanests}, while corresponding density plots of the objective function parameters are given in Figure~\ref{fig:compareModels} in the Appendix. Looking at the estimates for both data sets, a few patterns emerge. First, we can see the same relationship between the outdegree parameter, $\beta_1$, and the reciprocity parameter, $\beta_2$ in both data sets and across all three models. The estimates of $\beta_1$ are all negative and range from about -5 and -3, while the estimates of $\beta_2$ are all positive and range from about four to five. %In addition, the outegree parameter, $\beta_1$, \textit{decreases} when $\beta_3$ is included, while the reciprocity parameter, $\beta_2$ increases. This inverse relationship implies that the girls prefer to form small, closely knit friend groups, as opposed to being popular and having many friends. Tight-knit friend groups are more encouraged by model M2 than by models M1 and M3.
This suggests that in both data sets, nodes prefer outgoing ties that are reciprocated to those that are not. When we consider the context, this seems obvious: the students want to have close mutual friendships, and senators want mutual support for their bills. We are also concerned, however, with the extremely high correlation between the estimates of $\beta_1, \beta_2$. In all six model-data combinations, the absolute correlation between $\hat{\beta}_1, \hat{\beta}_2$ is greater than 0.93. %We explore the relationship between $\beta_1$ and $\beta_2$ further in Section~\ref{sec:samemodsamedat}.
% We have seen how the same model forms behave in different context, with some  differences and some similarities. 
After seeing the high correlation between parameter estimates in very different data contexts, we might question the guideline that both parameters need to be included in the objective function.

In Figure~\ref{fig:distests}, we saw that the inclusion of $\beta_3$ for the friendship data has a noticeable effect on the other parameters in the model. In Figure~\ref{fig:compareModels}, however, we see that this effect is not present in the senate data. Furthermore, the estimates of $\beta_4$ for the senate data are near zero, suggesting that this effect, which considers indirect ties, is not informative for the senate data. In contrast for the friendship data, the mean estimate of $\beta_4$ is about -7.5, which suggests that indirect ties are \textit{strongly} discouraged from forming. Again, this is fairly intuitive given the context: teenage girls likely want to have direct friendships with each other. Senators, on the other hand, have to wheel and deal to pass their legislation, and indirect relationships may be beneficial at times. By examining the results of models fitted to different data, we better understand the interpretation of the objective function parameters, and we have learned of some potential parameter correlation problems affecting the fit of the models. All in all, we have used four different ways of viewing collections of models, which have given us insight into the behavior of SAOMs.

@@

<<heatmap, fig.align='center', fig.cap='Adjacency matrix visualizations showing the theoretical (left) and empirical (right) transition probabilities for the first microstep observed in 1,000 simulations. The ego node is on the y-axis, and the alter node is on the x-axis. There are many ties with empirical probability zero.', fig.height=2.5>>=

@@

We use another seriated adjacency matrix visualization to show theoretical and empirical transition probabilities of the first microstep in Figure~\ref{fig:heatmap} for each tie that was the first change in one of the simulations. The figure is noticeably sparse: %showing the lack of coverage in the model.

>>>>

%Every good \emph{data} analysis includes both numerical and visual summaries of the data, so why restrict \emph{model} analysis, description, and diagnostics to numerical summaries? 
\emph{Model visualization} complements traditional model diagnostic tools such as $R^2$ values and residual plots to enrich our understanding and interpretation of statistical models. Each of the three different meanings of the word ``model", the model \emph{family}, the model \emph{form}, and the \emph{fitted} model \citep[see][pg. 2]{modelvis}, can benefit from model visualization (model-vis). We apply the three principles of model-vis: view the model in the data space, visualize collections of models, and explore the process of fitting the model, to the models and data introduced in Section~\ref{sec:exdatamods}. Since we have already decided on our model family (SAOMs), we shift our focus to the fitted model and the model form. Specifically, we use model-vis to learn more about how the \emph{model form} affects the \emph{fitted model}.

First, we put models in the data space to assess the model fit. %In \citeauthor{modelvis}, they chose define the \textit{data space} as ``the region over which we can reliably make inferences, usually a hypercube containing the data" \citep[p.~206]{modelvis}. But what does this definition mean for dynamic network models?
We define the \emph{data space} for SAOMs as the combined data structures of the network: the node, edge, and time information. We use the \texttt{R} package \texttt{geomnet} \citep{geomnet} to visualize the three data spaces together. The package lays out the network using node-link diagrams, and the network's node or edge data are then mapped to different visual features.  The color, size, and shape of the points can be used to represent variables in the node data, while the color, linewidth, and linetype of the edges between nodes can be used to represent variables in the edge data. To view temporal changes, we place the network at different timepoints side-by-side to see the evolution. By pulling all of this information together using \texttt{geomnet}, we can show the entire data space at once.

<<funsenatevis, fig.height=3, fig.cap='The 111$^{th}$ Senate at two times: while Clinton was in the senate in 2009 (left) and after she departed (right). We map sex, party, and bills authored to the shape, color, and size of the nodes, respectively. We also map $WPC$ to the edge width.'>>=
se111clint <- filter(seobama, senate == 111)
se111noclint <- filter(seobama, senate == 111)
se111noclint$target[which(se111noclint$target == "Hillary Rodham Clinton")] <- NA
se111clint$Clinton <- 'Yes'
se111noclint$Clinton <- 'No'
clintonSenate <- rbind(se111clint, se111noclint)
clintonSenate$Clinton <- as.factor(clintonSenate$Clinton)
clintonSenate$Clinton <- ordered(clintonSenate$Clinton, levels = c("Yes", "No"))

set.seed(12981120)
clintonSenate %>% filter(!(source %in% c("Roland Burris", "Bernard Sanders", "Frank R. Lautenberg", "Mary L. Landrieu") & is.na(target))) %>%
ggplot() +
  geom_net(directed = T, labelon=F, arrowsize = .3, singletons= F, fiteach = T, arrowgap = .01, layout.alg = 'fruchtermanreingold', ecolour = "grey70",
           aes(from_id = source, to_id = target, color = party, linewidth = gsw, shape = sex,size = n_au/30)) +
  scale_size_continuous(name="# bills\nauthored" , guide = guide_legend(nrow= 2), breaks = c(0,1,2,3), labels = c(0,30, 60, 90), range = c(.5,3)) +
  #theme_net() +
  ThemeNet +
  scale_shape_manual(values = c(17,16), guide = guide_legend(nrow= 2)) +
  scale_color_manual(values = c("black", "grey80","#ef3b2c"), guide = guide_legend(nrow=2)) +
  facet_wrap(~Clinton, nrow = 1, labeller = 'label_both')  +
  theme(legend.position = "bottom")
@

To demonstrate the functionality of \texttt{geomnet}, we visualize the connections in the $111^{th}$ U.S. Senate at two different timepoints: when Hillary Clinton was in the senate, and after she left to become Secretary of State. Clinton was only in the $111^{th}$ senate from January 3-20, 2009, in the middle of her second term as the junior senator from New York. In that time, she authored two bills and was a cosponsor on 17. There are many senators who cosponsored both of her bills, giving their edges to her a $WPC$ of 1. So with Clinton included in the node-link diagram in Figure~\ref{fig:funsenatevis}, the senate looks much more collaborative than it does without her. We then map our potential model covariates to visual features: number of bills authored to size of the node, gender to shape, and party to color. In addition, we visualize the strength of the tie by mapping the $WPC$ value between the two senators to the linewidth. By viewing node information, edge information, and time simultaneously, we have learned that Senator Clinton greatly affected the structure of collaboration within the senate.

<<summnetM1, message=FALSE, fig.cap='On the left, the first wave of observed data, conditioned on in the model. On the right, the second wave of observed data. In the middle, a summary network from the first model fit to the data. This summary network represents 1,000 simulations of wave 2 using the values from the simple fitted model M1.',results='hide', fig.height=2.5>>=
friendshipData <- array(c(fd2.w1, fd2.w2, fd2.w3), dim = c(16, 16, 3))
friendshipData <- array(c(fd2.w1, fd2.w2, fd2.w3), dim = c(16, 16, 3))
drink1 <- drink[20:35,1]
drink2 <- drink[20:35,2]
drink3 <- drink[20:35,3]
friendship <- sienaDependent(friendshipData)
alcohol <- varCovar(cbind(drink1,drink2, drink3))
mydata <- sienaDataCreate(friendship, alcohol)
M1eff <- getEffects(mydata)
M2eff <- includeEffects(M1eff, jumpXTransTrip, interaction1 = "alcohol")
M3eff <- includeEffects(M1eff, nbrDist2twice)
simu2 <- read.csv("data/simulation-1000-M1-M2-M3.csv")
means <- simu2 %>% group_by(Model, parameter) %>% summarize(
  means = mean(estimate)
)
M1parms <- (means %>% filter(Model == "M1"))$means
M2parms <- (means %>% filter(Model == "M2"))$means
M3parms <- (means %>% filter(Model == "M3"))$means
set.seed(4231352)
#M1sims1000 <- netvizinf::saom_simulate(dat = mydata,
#                            struct = M1eff,
#                       parms = as.numeric(M1parms),
#                       N = 1000)
#save(M1sims1000, file = "data/M1sims1000.RData")

@@

To view a SAOM in the data space, we simulate from the fitted model. No single network instance simulated from a SAOM will fully represent the model, so instead we look at an \textit{expectation} from the network model. We frequently rely on expected values in statistics, but network models, especially those as complex as SAOMs, lack an expected network measure. We could talk about expected values of parameters, but the parameters can be hard to interpret and cannot teach us much about the overall structure of the network. How then, can we arrive at an ``expected" network? We answer this question with visualization.
%Expected values of parameters can be computed, but if they cannot directly tell us anything about our dependent variable, they lose value. 
%There is no defined expected value of a network model, though expectations exist for most other statistical models. 
%For network data, we view an ``average" network through a sort of summary network drawn using the traditional node-link diagram. 
%In Figure~\ref{fig:summnetM1}, we show the \textit{expected network} of wave 2 from model M1, created with 1,000 simulations from the model. 
To construct an expected network, we
\begin{enumerate}
\item simulate 1,000 wave 2 and wave 3 instances of our small friendship network from model M1 using estimated parameter values (see Table~\ref{tab:meanests}) that have previously been estimated. Then, we 
\item count the number of times each edge $i \rightarrow j$ appeared in one of the 1,000 network simulations. Then, we
\item construct one weighted summary network for time points 2 and 3. Each summary network has edge weights equal to the proportion of the simulated networks in which the edge appeared. Finally, we 
\item create the \textit{expected network} by including only the edges that appeared in more than 5\% of the simulations. 
\end{enumerate}
The resulting expected network is in the middle panel of Figure~\ref{fig:summnetM1}. Edges with higher weights are darker, thicker lines. On either side of the average network in Figure~\ref{fig:summnetM1}, we show the data for comparison: wave 1 on the left, and wave 2 on the right. We can see that the overall structure of the average network is much more similar to wave 1 than to wave 2. In both wave 1 and the expected network, the edges form two distinct clusters of nodes: $\{2,3,4,5,6,12,13,15\}$ and $\{7,8,9,10,11,14\}$. The expected network, however, is supposed to represent the second wave of data, on the right in Figure~\ref{fig:summnetM1}. The expected network does not resemble wave 2, which has one small, strongly connected cluster and one larger, sparsely connected cluster, with four unconnected nodes. Because the expected network does not resemble wave 2, we take this as evidence that the simple model, M1, does not capture the change in network structure from the first to the second wave of data. Visualizing the expected network in this way places the model M1 in the data space, and helps us assess model fit. % We can use the average network to help determine model goodness-of-fit. Because the the expected  network looks more like the first wave than the second wave, we can use the visualization in Figure~\ref{fig:summnetM1} as evidence of poor model fit.

@@

Another way to view a SAOM in the data space is with \textit{difference networks} between the expected network and the data, as shown in Figure~\ref{fig:summary2}. We created these networks by computing the difference between the adjacency matrices for the expected network and wave 1, and for wave 2 and the expected network. The lack of model fit becomes even more clear in Figure~\ref{fig:summary2}, with the broken edges representing ties that only show up in the expected network, and the solid edges representing ties in the data that the model fails to capture.  Dotted edges are in both the expectation and the data and represent where the model ``got it right". Again, the widths of the edges demonstrate the frequency of the ties in the simulations. On the left in Figure~\ref{fig:summary2}, we see the difference between the first wave of data and the expected network. We only see dotted and broken edges, meaning that the simulated networks consist primarily of edges in wave 1, plus some new edges that are not in the first wave of data. The unobserved ties are all reciprocating edges in the data, so the model predicts more reciprocity between actors than there actually is. On the right in Figure~\ref{fig:summary2}, we see the difference between the second wave of data and the expected network. Because the model is simulating the second wave of data, we would hope to see very few differences between the expected network and wave 2. However, there are six solid edges, which are edges in the data that do not appear at all in the expected network, and there are many broken edges that appear in the expected network but not in the data. Although there are some edges that appear in both the expected network and wave 2, the majority of edges in the expected network are not in wave 2, and there are several edges in the data missing from the expectation. We conclude that the model M1 is a poor fit to our example data, which we have seen very clearly by viewing the model in the data space.

@@

Now that we have assessed model fit, we want to learn about model behavior in order to choose better fitting models in the future. This is the second principle of model visualization: ``collections [of models] are more informative" \citep{modelvis}. We examine the behavior of SAOMs by visualizing collections of them, which we construct in four ways. First, we explore the space of all possible models to determine the most significant model parameters.  Second, we vary model settings to uncover relationships between parameters and their effects on the fitted model. Next, we fit the same model to the same data many times to assess the behavior of the underlying simulations. Finally, we fit the same model to different data to learn about fitted parameter values under different conditions. We chose these four collections because they each explore something different about SAOM behavior: the choice of parameters in the objective function, the effects of underlying simulations, and the resulting fitted values under different circumstances. %Using the tools in our visualization toolbox, we view these collections to learn about model behavior.%The first considers the many parameters to include in a SAOM, while the second results in different parameter estimates, so we can see how parameter estimates vary. The third collection shows how those many possible parameters affect the values of parameters in the fitted models. Finally, we explore how the same model behaves under two completely different circumstances.

@@

\noindent \textbf{Exploring the space of all possible models:} The \texttt{RSiena} manual contains over 80 effects to include in the SAOM objective function. To select objective function parameters for modeling our example data, we searched through all effects available to find \textit{significant} effects. The \texttt{RSiena} package includes the function \texttt{Wald.Rsiena} for Wald-type tests of parameter significance. We start by including the outdegree and reciprocity parameters in the model then add one effect to the objective function, fit the model to the example data, test the added effect for significance, and repeat for all possible model parameters. We performed this procedure, which we used to pick M2 and M3, for both the small friendship and the senate collaboration data. A visualization of the significant effects for the two example data sets are shown in Figure~\ref{fig:sigtesting} in the Appendix. Using this one-at-a-time testing method results in many possible significant parameters of different type and magnitude, demonstrating the flexibility of these models.  %We see in both the friendship data and the senate data results that most of the significant effects have absolute value less than ten. In addition, the $p$-values for the effects from the friendship data are more spread out than the $p$-values for the senate data, which are concentrated at about 0.02 or less. This may suggest that larger data sets tend to result generally in smaller $p$-values, just like a larger sample size results in smaller $p$-values in a $t$-test.

@@

\noindent \textbf{Varying model settings:} The second method for viewing collections of models is varying model settings. We have done this already by choosing three different models M1, M2, M3 to fit to our example data sets. To explore the behavior of these three models, we simulate from them. Using the estimated parameter values in Table~\ref{tab:meanests}, we simulated 1,000 observations of waves 2 and 3 ofthe friendship data from each of the three models. From the simulations, we created an expected network for wave 2 from models M1, M2, and M3 using the same method described in Section~\ref{sec:minds}. These expected networks are shown in Figure~\ref{fig:averagenet} in the Appendix, and we see that all models are failing to capture the structure of the second wave, even though the individual parameters were all significant. %In Section~\ref{sec:samemodsamedat}, we fit these three models to the data 1,000 times, and examine the parameter densities. Here, . For simulation, we used the mean parameter values from the 1,000 fits of the corresponding model.

@@

<<distests, fig.cap="Distribution of fitted parameter values for our three SAOMs. The inclusion of $\\beta_3$ or $\\beta_4$ clearly has an effect on the distributions of the rate parameters, $\\alpha_1$ and $\\alpha_2$, and on the other objective function parameters $\\beta_1, \\beta_2$.", fig.height=3>>=

@@

\noindent \textbf{Fitting the same model to the same data:} The simulations required to fit a SAOM lead to inherent variability in the fitted values. To examine this variability, we fit models M1, M2, and M3 to the friendship network 1,000 times each. The distributions of the fitted model parameters are shown in Figure~\ref{fig:distests}. Clearly, the inclusion of the JTT parameter, $\beta_3$, results in very different estimates the of the other four parameters in all models, $\alpha_1$, $\alpha_2$, $\beta_1$, and $\beta_2$. On the other hand, the inclusion of the indirect tie parameter, $\beta_4$, does not lead to such drastically different estimates of other model parameters. When $\beta_3$ is included, its estimate is postitive, meaning that friendships between two girls with different drinking behaviors tend to form when there is third girl who is already friends with the other two. The inclusion of $\beta_3$ also leads to increases in the estimates of $\alpha_1, \alpha_2$. Thus, when modeling the transitive triplet behavior in this way, we may conclude that girls change friends more frequently. It is not clear, however, that the addition of a parameter to the objective function \textit{should} effect the estimates of the rate parameters. This strong and possibly problematic relationship between rate parameter estimates and objective function parameters is further shown in Figure~\ref{fig:corplots} in the Appendix. When visualizing the results of the same models fit many times to the same data, we can to see potential model behavior problems right away.   %In comparison with models M1 and M3, model M2 typically has higher estimates of the rate parameter, meaning that the inclusion of the covariate statistic in the model leads to higher estimates of the number of times, on average, a node gets to change its ties.

@@

% table of fitted M1 M2 M3
\begin{table}
\scalebox{0.9}{
\begin{tabular}{lccc|ccc }

@@

\noindent \textbf{Fitting the same model to different data:} Finally, we view collections of models by fitting the same models to different data to understand how the model behaves in different contexts. As we did for the friendship data, we estimated the parameters for M1, M2, and M3, when fit to the senate data (see Figure \ref{fig:senateplot}) 100 times each. The means and standard deviations of the repeated parameter estimates for all six model-data combinations are given in Table~\ref{tab:meanests}, while corresponding density plots of the objective function parameters are given in Figure~\ref{fig:compareModels} in the Appendix. Looking at the estimates for both data sets, a few patterns emerge. First, we can see the same relationship between the outdegree parameter, $\beta_1$ and the reciprocity parameter, $\beta_2$ in both data sets and across all three models. The estimates of $\beta_1$ are all negative and range from about -5 and -3, while the estimates of $\beta_2$ are all positive and range from about four to five. %In addition, the outegree parameter, $\beta_1$, \textit{decreases} when $\beta_3$ is included, while the reciprocity parameter, $\beta_2$ increases. This inverse relationship implies that the girls prefer to form small, closely knit friend groups, as opposed to being popular and having many friends. Tight-knit friend groups are more encouraged by model M2 than by models M1 and M3.
This suggests that in both data sets, nodes prefer outgoing ties that are reciprocated to those that are not. When we consider the context, this seems obvious: the students want to have close mutual friendships, and senators want mutual support for their bills. We are also concerned, however, with the extremely high correlation between the estimates of $\beta_1, \beta_2$. In all six model-data combinations, the absolute correlation between $\hat{\beta}_1, \hat{\beta}_2$ is greater than 0.93. %We explore the relationship between $\beta_1$ and $\beta_2$ further in Section~\ref{sec:samemodsamedat}.
% We have seen how the same model forms behave in different context, with some  differences and some similarities. 
After seeing the high correlation between parameter estimates in very different data contexts, we might question the guideline that both parameters need to be included in the objective function.

In Figure~\ref{fig:distests}, we saw that the inclusion of $\beta_3$ for the friendship data has a noticeable effect on the other parameters in the model. In Figure~\ref{fig:compareModels}, however, we see that this effect is not present in the senate data. Furthermore, the estimates of $\beta_4$ for the senate data are near zero, suggesting that this effect, which considers indirect ties, is not informative for the senate data. In contrast, for the friendship data the mean estimate of $\beta_4$ is about -7.5, which suggests that indirect ties are \textit{strongly} discouraged from forming. Again, this is fairly intuitive given the context: teenage girls likely want to have direct friendships with each other. Senators, on the other hand, have to wheel and deal to pass their legislation, and indirect relationships may be beneficial. By examining the results of models fitted to different data, we better understand the interpretation of the objective function parameters, and we have learned of some potential parameter correlation problems affecting the fit of the models. All in all, we have used four different ways of viewing collections of models, which have given us insight into the behavior of SAOMs.

@@

<<heatmap, fig.align='center', fig.cap='Adjacency matrix visualizations showing the theoretical (left) and empirical (right) transition probabilities for the first microstep taken in 1,000 simulations. The ego node is on the y-axis, and the alter node is on the x-axis. There are many ties with empirical probability zero.', fig.height=2.5>>=

@@

We use another seriated adjacency matrix visualization to show theoretical and empirical transition probabilities of the first microstep in Figure~\ref{fig:heatmap} for each tie that was the first change in a simulation. The figure is noticeably sparse: %showing the lack of coverage in the model.
